% Pass options to the xcolor package for table support
\PassOptionsToPackage{table}{xcolor}

% Document class: A4 paper with an 6pt font size.
\documentclass[a4paper]{article}

% Use a plain page style (no headers/footers beyond page numbers)
\pagestyle{plain}

% Import the base preamble (packages, macros, environments, etc.)
\input{sections/base}

% Set the title and author metadata
\title{Algorithmic Trading}
\author{Arthur Li}

\begin{document}

% Use Roman page numbering for front matter
\pagenumbering{roman}

% Title Page
\maketitle

% Bibliography style (Chicago)
\bibliographystyle{chicago}

% Table of Contents
\newpage
\tableofcontents

% Switch to Arabic numbering for main content pages
\newpage
\pagenumbering{arabic}

% Introduction Section
\newpage
\input{sections/intro/overallstruct}

% Introduction Section
\newpage
\input{sections/global_macro/00_overallstruct}

% Research
\newpage
\input{sections/research/overallstruct}   

\begin{comment}
	
% Mathematical and Statistical Foundation
    
    Adapted from materials such as Shreve’s stochastic calculus texts and advanced statistics references)

Stochastic Processes and Bayesian Inference:
Topics: Stochastic calculus, Markov processes, Bayesian linear regression, MCMC methods.
Purpose: Equip readers with rigorous tools to model market randomness and update beliefs with new data.
Optimization Techniques:
Topics: Convex optimization, dynamic programming, optimal control theory.
Purpose: Provide the mathematical underpinnings for optimal execution and risk management.
	
	% Quantitative Finance Theory
	(Incorporating insights from market microstructure texts and execution research)

Market Microstructure and Optimal Execution:
Topics: Order flow, transaction cost analysis, smart order routing, and alpha profiling.
Example: Discuss methods like VWAP, TWAP, and implementations from scholarly articles and Wikipedia’s smart order routing overview 
EN.WIKIPEDIA.ORG
.
Derivative Pricing and Risk Management:
Topics: Option pricing models, risk-adjusted cost functions, and sensitivity analysis.
Purpose: Bridge the gap between theoretical pricing models and their algorithmic application.

	% Algorithmic Trading Strategies
	(Drawing on works such as Michael Halls Moore’s “Advanced Algorithmic Trading” 
SACREDTRADERS.COM
and Ernie Chan’s insights)

Strategy Design and Testing:
Mean reversion, momentum, delta-neutral and statistical arbitrage strategies.
Detailed discussion of backtesting pitfalls and strategy decay.
Execution Algorithms:
In-depth treatment of execution strategies including slicing orders, smart order routing, and adaptive algorithms.
Case Studies:
Real-world examples (e.g., pairs trading with cointegration and Kalman filtering) that show the evolution from theoretical model to live trading.

	% Machine Learning and AI in Trading
	(Incorporating insights from “Advances in Financial Machine Learning” by Lopez de Prado
ANALYZINGALPHA.COM
 and related texts)

Supervised and Unsupervised Techniques:
Regression, decision trees, clustering, and support vector machines applied to financial data.
Deep Learning and Reinforcement Learning:
Neural networks for pattern recognition and forecasting, including practical coding examples in Python.
Feature Engineering and Model Evaluation:
Techniques for constructing and validating predictive models, with an emphasis on avoiding overfitting in a trading context.

	% Implementation, Backtesting, and Deployment
	(Using open-source tools like QSTrader, Python, and R – as highlighted in QuantInsti’s advanced courses 
QUANTRA.QUANTINSTI.COM
)

Efficient Backtesting Techniques:
Vectorized and event-driven backtesting frameworks.
Risk and Performance Metrics:
Detailed analysis of performance, risk-adjusted returns, and practical considerations when moving to live execution.
Real-Time Deployment:
Integrating machine learning models with live data streams, execution engines, and risk management systems.

	% Advanced Topics and Future Directions
	Emerging Trends:
Discussion on topics such as high-frequency trading (HFT) infrastructure, smart order routing, and dark pool strategies.
Reference recent research and articles (e.g., the “mundane economics” of HFT ).
Regulatory and Market Implications:
How evolving market regulations impact algorithm design and execution.

	% Appendices and Supplemental Materials
	Mathematical Proofs and Derivations:
Detailed proofs for key results for those interested in the theory.
Code Libraries and Examples:
Fully annotated code snippets (Python/R) that implement the discussed algorithms.
Glossary and Further Reading:
A curated list of additional resources and advanced texts (referencing lists from QuantInsti and other curated guides 
BLOG.QUANTINSTI.COM
).

    % Mathematical Overview Section
    \newpage
    %\input{sections/math/overallstruct}

    % Microstructure Trading Section
    \newpage
    %\input{sections/microstruct/overallstruct}

    % Equities Trading Section
    \newpage
    %\input{sections/equities/eqtrade}

    % Fixed Income Trading Section
    \newpage
    %\input{sections/fixedincome/fitrade}

    % Derivatives Trading Section
    \newpage
    %\input{sections/derivatives/dertrade}

    % Forex Trading Section
    \newpage
    %\input{sections/forex/forextrade}

    % Commodities Trading Section
    \newpage
    %\input{sections/commodities/commotrade}

    % Appendix Section
    \newpage
    %\input{sections/appendix/appindex}
 
\end{comment}

\newpage
\bibliography{references}

\end{document}
	